A Generalized Visbility to Power-Spectrum Framework for 21\,cm Inteferometric Probes of Reionizations

### 2017-07-02 Rotation paper/draft_preWterm.tex
save

%\documentclass[preprint2,numberedappendix,tighten,twocolappendix]{aastex6}  % USE THIS TO MAKE BIB, THEN FORMAT USING EMULATEAPJ
\documentclass[twocolumn,apj,numberedappendix]{emulateapj}
\shorttitle{Adding Sensitivity to 21\,cm Inteferometric Probes of Reionization}
\shortauthors{Zhang\&Parsons}
\usepackage{float}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[figuresright]{rotating}
%\usepackage{rotating}
\usepackage{natbib}
%\usepackage{pdflscape}
%\usepackage{lscape}
\usepackage{ctable}
\citestyle{aa}
\renewcommand\[{\begin{equation}}
\renewcommand\]{\end{equation}}
\graphicspath{{../figures/}}

@@

\title{Adding Sensitivity to 21\,cm Inteferometric Probes of Reionization by Optimizing Choice of Baselines}

@@

Arguably the most promising observational probe of the epoch of reionization
comes from measurement of the ``spin-flip'' transition of neutral
hydrogen of characteristic wavelength 21\,cm \citep{Furlanetto2006181,PritchardLoeb}.
Other observational studies of reionization include for example Gunn-Peterson measurements of quasi-stellar objects \citep{Fan2006}. GP tests rely on Ly$\alpha$ absorption, and thus easily saturate with small degree of ionization and can only probe the end stage of reionization. [other probes include cmb optical depth, cmb ksz, lyman alpha emitters (redshifted as in bubble to no longer be absorbed, big bubbles corresponding to clustered sources), GP, UV luminosity of high redshift galaxy] The 21\,cm brightness temperature distribution is a direct tracer of neutral hydrogen through the epoch of reionization, and thus measurements of the 21\,cm line would potentially
shed light on all stages of reionization. Radio interferometric efforts
to measure the 21\,cm power spectrum [spatial pspec of brightness temp fluctuations] has been a top priority in recent years of astronomy.
Current-generation instruments include the Precision Array for Probing
the Epoch of Reionization (PAPER) \citep{Ali2015,paper32}, Murchison
Widefield Array (MWA) \citep{Bowman2013, Tingay2013}, Low Frequency Array (LOFAR) \citep{LOFAR}. Next-generation instruments include the Hydrogen Epoch of Reionization
Array (HERA) (e.g. \citealt{HERA,HERAconfiguration,HERABEAM1,HERADISH2})  under construction, 
and the Square Kilometer Array Low (SKA-low) (e.g. \citealt{SKA1}), which is currently in planning.

The highly redshifted 21\,cm signal is faint and diffuse, in contrast to the localized bright sources that have been of interest in many traditional applications of radio astronomy. Thus modern low-frequency radio interferometers aimed at measuring the 21\,cm signal bears designs different from traditional instruments.  Current experiments are sensitivity-starved, with the sensitivity requirement further increased considering the foreground contaminations 5 orders of manitude brighter. With the goal of detecting a cosmological signal 5 orders of magnitude fainter than foreground contaminations, current efforts are sensitivity starved. The arrays are thus designed with multiple copies of the same baselines to repeatedly measure the same Fourier signal to increase sensitivity. To satisfy the sensitivity need, modern arrays are large. Ranging upwards from 100 elements, these arrays are typically of the drift-scan type to limit cost.

Analysis pipelines for these new arrays typically fall into two categories. In the first, more traditional technique, images are formed in Fourier domain through rotation synthesis, followed by a foreground mitigation step to construct a power spectrum. An alternative technique works directly with visibilities from baselines and cross multiplies them to form the power spectrum. Tracking closely the instrumental output of visibilities, this technique allows one to deal with systematics such as Radio-Frequency-Interference(RFI) and instrumental cross-talks in a more native context, and thus tend to provide good limits on the power spectrum. An example of the visibility-based pipeline was presented in \cite{Ali2015}, which provides the newest upper limit to the power spectrum
measurements with the 64-element version of PAPER (henceforth as PAPER-64). However, one disadvantage of existing visibility-based pipelines is the lack of use of partial redundancy. Baselines of different length and orientation still contain partially redundant information. While an imaging based power-spectrum pipeline naturally includes all redundancy information, visibility-based pipelines until now only cross multiply fully redundant baselines, in other words baselines of the same length and orientation.

@@

[TODO: this paragraph would potentially be promoted to a section] By using pairs of baselines, we also reduce additional mode mixing. As discussed in \cite{Hazelton2013}, the varying shapes of overlapping regions in $uv$ space of all contributing baselines lead to chromaticities in the response. By considering only pairs of baselines, we fix shapes of overlapping regions, and only let their sizes vary with frequency, just as in the case of fully redundant baselines.

@@

\caption{The PAPER-128 layout. Each dot corresponds to the location of
an antenna. Top panel shows the antenna positions drawn to scale;
bottom panel show the antenna labels and distances of the 112-element grid, in other words excluding the outrigger antennas.
The numbering of the antennas in the bottom panel are original labels
during instrument assembly and does not bear significant
meaning. In the bottom panel, the two baselines (5,77) and (40,85) marked with red segments are example of an equivalent pair, both with separation denoted by \{1, 0\}, for the
antennas are separated by 1 unit east and 0 unit north. Similarly,
the baselines marked in blue and yellow are examples
of the classes \{1, 1\} and \{1,-1\}, respectively. Due to the small North-South separation within the grid, \{1, 1\} is expected to be near-equivalent to \{1, 0\}. }
\label{fig:AntPos}
\end{figure}

We shall use the 128-element PAPER array (henceforth as PAPER-128) to motivate our formalism and demonstrate our method, and extend our results to several HERA configurations in Sec \ref{sec:arrconf}. 
The PAPER array is located in the Karoo desert in South Africa (30:43:17.5
S, 21:25:41.8 E). The layout pattern with antenna labels are show
in Fig. \ref{fig:AntPos}. The array consists of a 112-element grid and 16 "outriggers" used primarily to aid calibration.  In the bottom panel, the two baselines (5, 77) and (40, 85), marked in red, are an example of equivalent pair. We denote a equivalency-class of baselines in the PAPER grid by their separations, in this case  \{1, 0\}, for the
antennas are separated by 1 unit east and 0 unit north. Similarly,
the baselines marked in blue and yellow are respectively examples
of \{1, 1\} and \{1, -1\}.
Note \{1, 0\} and \{-1, 0\} for example are the same class and should
not be counted twice. Antennas in purely north-south baselines
are close (4m), and hence these baselines are not suitable
for cross-multiplication due to cross-talks. On the other hand, the close North-South separation means that classes such as \{1, 0\} and  \{1, 1\} are expected to be near-equivalent. The PAPER-64 analysis of \cite{Ali2015} used three classes of baselines, the PAPER-128
equivalent of which are 
\{2, 0\}, \{2, 1\} and \{2, -1\}. There each of these classes
of baselines are cross multiplied within itself. This paper provides the method for inter-class multiplications.

@@

\subsection{Rough Idea}

Given a point source on the sky, each baseline maps the
source to a point in $uv$ plane. As the Earth rotates with respect to
the source, the point traces out tracks in the $uv$ plane. 
We show in Fig. \ref{fig:Tracks} $uv$ tracks of PAPER-128 over 4.8 hours, at 0.15GHz, for a source that passes through zenith.

@@

Roughly speaking, we can identify
redundancy of near-equivalent baselines as crossings
of the $uv$ tracks. As we see in Fig. \ref{fig:Tracks}, there are many such crossings. Crossing tracks, however does not imply perfect redundancy. The reason lies in the finite size of the beams. We show sample beams of HERA and PAPER antennas in Fig. \ref{fig:Beam} for reference. The inaccuracy is expected to be even more significant for the smaller HERA beam, which has a larger corresponding $uv$ point spread function.

Thus although it is a valid quick method to identify some baselines to cross-multiply, track-crossing is
not accurate enough for time offset determination, nor can give estimate of the degree of redundancy. To determine the time-offset corresponding to maximal redundancy, we must develop a more general  formalism that accounts for the point spread function of the finite beams, and estimates the degree of redundancy for general combination of baselines at a general time-offset, as we do in the next section.

@@

\caption{Sample beam response of HERA (left) and PAPER (right) antennas, both
for frequency of $\nu=150\text{MHz}$ and Stokes $I$ polarization. Notice $A$ in this paper is a ``baseline's beam'', equivalent to squares of the antenna voltage beams shown here. The circles centered around zenith (center of beam) here are
spaced 10 degrees apart. The finite size of the beams limits the accuracy of the $uv$ track-crossing as a method of redundancy search. \label{fig:Beam}}
\end{figure}

\subsection{Formalism}
Below we shall derive theoretical expectations of cross multiplications
of two near-equivalent baselines. More precisely we shall relate
the product of visibilities of two different baselines to the power spectrum.

We take the visibility as commonly defined in the literature (e.g.
\citealt{first-paper}): 
\begin{equation}
\begin{aligned}V_{\nu}(\boldsymbol{b}) & =\int d\Omega A_{\nu}(\hat{\boldsymbol{s}})\phi(\nu)I_{\nu}(\hat{\boldsymbol{s}})\exp\left[-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right],\\
 & \approx\frac{2k_{B}}{\lambda^{2}}\int d\Omega A_{\nu}(\hat{\boldsymbol{s}})\phi(\nu)T(\hat{\boldsymbol{s}})\exp\left[-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right],
\end{aligned}
\label{eq:Vis1}
\end{equation}
Here $\lambda$ is a mean wavelength, $\boldsymbol{b}$ is the baseline
length, $\hat{\boldsymbol{s}}$ and $\Omega$ are a direction in the
sky and its corresponding solid angle. $A_{\nu}$ is the (frequency
dependent) beam, and $I$ is the specific intensity, which has been
related to $T$, the brightness temperature in the Rayleigh-Jeans
limit. $k_B$ is Boltzmann's constant. Note that $A$ here is the baseline beam, i.e. product of two antenna voltage beams, examples of which are shown in Fig. \ref{fig:Beam}. $\phi(\nu)$ is the frequency bandpass profile. In practice,
power spectrum measurements are typically taken from a few ten MHz centered around the corresponding redshift of interest (e.g. 150 MHz for z=9.5). We take a moment to note that in Eq. \ref{eq:Vis1}, no flat-sky approximation has yet been made. The angular integral is performed over the dome $d\Omega$.

We define the delay-transformed visibility \citep{delay-transform}:
\small
\begin{equation}
\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left[-2\pi i\nu\tau\right],\\
 & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{\boldsymbol{s}},\nu)\phi(\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}+\tau\right)\right]
\end{aligned}
.\label{eq:Vb1}
\end{equation}
\normalsize

Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
an integral over observation coordinates $\hat{\boldsymbol{s}}$ and $\nu$. Ultimately,
we would like to relate the data, collected with coordinates $\hat{\boldsymbol{s}}$
and $\nu$, to the power spectrum, written with cosmological coordinates
$\boldsymbol{r}$ and $\boldsymbol{k}$. We start by noticing that
\[
\begin{aligned}r & =\frac{c}{H_{0}}\int_{0}^{z}\frac{dz'}{E(z')},\\
 & \approx\frac{c}{H_{0}}\int_{0}^{z_{0}}\frac{dz'}{E(z')}-\frac{c(1+z)^{2}}{\nu_{21}H_{0}E(z)}\left(\nu-\nu_{0}\right),\\
 & \equiv D_{c}-Y\Delta\nu,
\end{aligned}
\]
where $\nu_{21}=1420$MHz is the 21\,cm transition rest frequency, $\nu_{0}$
a reference central frequency with corresponding redshift $z_{0}$,
and 
\[
E(z)=\sqrt{\Omega_{m}(1+z)^{3}+\Omega_{\Lambda}}.
\]
Inverting for $\nu$:
\begin{equation}
\nu=\frac{D_{c}-r}{Y}+\nu_{21}.\label{eq:nur}
\end{equation}

Further we shall invoke the flat-sky limit
\begin{equation}
(r_{x},r_{y})=(X\hat{\boldsymbol{s}}_{x},X\hat{\boldsymbol{s}}_{y}), 
\end{equation}
where
\begin{equation}
X = \frac{c}{H_0}\int_{0}^{z}\frac{dz'}{E(z')}
\end{equation}
is a cosmological scalar. In the same limit $(Xk_{x},Xk_{y},Yk_{z})=\frac{2\pi}{c}(b_{x},b_{y},\tau)$ relates
the cosmological $k$ to the observation coordinates. For a more rigorous treatment without the flat sky limit, see Appendix [ADRIAN].

We can rewrite the delayed transformed visibility as 
\small
\[
\begin{aligned}V(\boldsymbol{b},\tau) & =\frac{2k_{B}}{\lambda^{2}}\int_{H}\frac{d^{3}r}{X^{2}Y}A(\boldsymbol{r})\phi(r)T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}+\tau\right)\nu_{r}\right]\end{aligned},
\]
\normalsize
 where $d\nu=-dr/Y$ and $d^{3}r=-X^{2}Yd\Omega d\nu$.

We have written $\nu_{r}$ to remind us that $\nu$ and $r$ are related
by Eq. \eqref{eq:nur}. The beam reception pattern $A$ is dimensionless,
normalized to 1 at its peak (zenith), and we assume it to be the same
for all baselines.

With a time offset, the beam pattern has moved relative to the sky.
Here we choose to fix the sky, and denote the rotated coordinates
of the beam pattern with the 3 dimensional rotation operator $\Gamma$. With implicit bounds
of integrals from $-\infty$ to $\infty$, we have:
\begin{widetext}
\begin{equation}
\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau')\rangle\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & \approx\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rd^{3}r'\delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}r|A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r})||\phi(\nu_{r})|^{2}\exp\left[-i2\pi\nu_{r}\left(\hat{\boldsymbol{r}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma \hat{\boldsymbol{r}}\cdot\frac{\boldsymbol{b'}}{c}\right)\right],\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma \hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2}\exp\left[-i2\pi\nu\left(\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b}}{c}-\Gamma\hat{\boldsymbol{s}}\cdot\frac{\boldsymbol{b'}}{c}\right)\right],
\end{aligned}
\label{eq:main}
\end{equation}

>>>>

%\documentclass[preprint2,numberedappendix,tighten,twocolappendix]{aastex6}  % USE THIS TO MAKE BIB, THEN FORMAT USING EMULATEAPJ
\documentclass[twocolumn,apj,numberedappendix]{emulateapj}
\shorttitle{Adding Sensitivity to 21\,cm Inteferometric Probes of Reionization}
\shortauthors{Zhang, Liu \&Parsons}
\usepackage{float}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[figuresright]{rotating}
%\usepackage{rotating}
\usepackage{natbib}
%\usepackage{pdflscape}
%\usepackage{lscape}
\usepackage{ctable}
\citestyle{aa}
\renewcommand\[{\begin{equation}}
\renewcommand\]{\end{equation}}
\graphicspath{{../figures/}}

@@

\title{A Generalized Visbility to Power-Spectrum Framework for 21\,cm Inteferometric Probes of Reionizations}

@@

In the past, observational studies of reionization, including Gunn-Peterson measurements of quasi-stellar objects \citep{Fan2006} and cosmic microwave background (CMB) measurements of temperature and anisotropy \citep{Planck2016}, Kinetic Sunyaev-Zeldo'vich effect \citep{kszpatchy} and and Lyman alpha emitter clustering \citep{mcquinnLyA} have given us indications of the rough time-frame of reionization, but only limited constrains on the finer spatial and temporal variations. A surge of recent radio-astronomical experiments of reionization focus on measurement of the ``spin-flip'' transition of neutral
hydrogen of characteristic wavelength 21\,cm \citep{Furlanetto2006181,PritchardLoeb}.
In complement to the aforementioned probes, the 21\,cm brightness temperature distribution is a direct tracer of neutral hydrogen through the epoch of reionization, and thus tomography of the 21\,cm line is a direct measurement of the full temporal and spatial variations of this event. 
Before possible realization of full scale 21\,cm tomography, many current radio interferometric efforts
prioritize measuring the spatial power spectrum of 21\,cm brightness temperature fluctuations.
Current-generation instruments include the Precision Array for Probing
the Epoch of Reionization (PAPER) \citep{Ali2015,paper32}, Murchison
Widefield Array (MWA) \citep{Bowman2013, Tingay2013}, Low Frequency Array (LOFAR) \citep{LOFAR}. Next-generation instruments include the Hydrogen Epoch of Reionization
Array (HERA) (e.g. \citealt{HERA,HERAconfiguration,HERABEAM1,HERADISH2})  under construction, 
and the Square Kilometer Array Low (SKA-low) (e.g. \citealt{SKA1}), which is currently in planning.

The highly redshifted 21\,cm signal is faint and diffuse, in contrast to the localized bright sources that have been of interest in many traditional applications of radio astronomy.  Current experiments are sensitivity-starved, with the sensitivity requirement further increased considering the foreground contaminations 5 orders of magnitude brighter than the cosmological signal of interest. Thus modern low-frequency radio interferometers aimed at measuring the 21\,cm signal bears designs different from traditional instruments. Many specially designed arrays, such as PAPER and HERA, feature multiple copies of the same baselines to repeatedly measure the same Fourier signal to increase sensitivity. To satisfy the sensitivity need, modern arrays are large. Ranging upwards from 100 elements, these arrays are typically of the drift-scan type to limit cost.

Analysis pipelines for the 21\,cm power spectrum typically fall into two categories. In the first, more traditional technique, images are formed in Fourier domain through rotation synthesis, followed by a foreground mitigation step to construct a power spectrum. An alternative technique works directly with visibilities from baselines and cross multiplies them to form the power spectrum. Tracking closely the instrumental output of visibilities, this technique allows one to deal with systematics such as Radio-Frequency-Interference(RFI) and instrumental cross-talks in a more native context, and thus tend to provide good limits on the power spectrum. An example of the visibility-based pipeline was presented in \cite{Ali2015}, which provides the newest upper limit to the power spectrum
measurements with the 64-element version of PAPER (henceforth as PAPER-64). However, one disadvantage of existing visibility-based pipelines is the lack of use of partial redundancy. Baselines of different length and orientation still contain partially redundant information. While an imaging based power-spectrum pipeline naturally includes all redundancy information, visibility-based pipelines until now only cross multiply fully redundant baselines, in other words baselines of the same length and orientation.

@@

%[TODO: this paragraph would potentially be promoted to a section] By using pairs of baselines, we also reduce additional mode mixing. As discussed in \cite{Hazelton2013}, the varying shapes of overlapping regions in $uv$ space of all contributing baselines lead to chromaticities in the response. By considering only pairs of baselines, we fix shapes of overlapping regions, and only let their sizes vary with frequency, just as in the case of fully redundant baselines.

@@

\caption{The PAPER-128 layout. Each dot corresponds to the location of
an antenna. Top panel shows the antenna positions drawn to scale;
bottom panel show the antenna labels and distances of the 112-element grid, in other words excluding the outrigger antennas.
The numbering of the antennas in the bottom panel are original labels
during instrument assembly and does not bear significant
meaning. In the bottom panel, the two baselines (5,77) and (40,85) marked with red segments are example of an equivalent pair, both with separation denoted by \{1,0\}, for the
antennas are separated by 1 unit east and 0 unit north. Similarly,
the baselines marked in blue and yellow are examples
of the classes \{1,1\} and \{1,-1\}, respectively. Due to the small North-South separation within the grid, \{1,1\} is expected to be near-equivalent to \{1,0\}. }
\label{fig:AntPos}
\end{figure}

We shall use the 128-element PAPER array (henceforth as PAPER-128) to motivate our formalism and demonstrate our method, and extend our results to several HERA configurations in Sec \ref{sec:arrconf}. 
The PAPER array is located in the Karoo desert in South Africa (30:43:17.5
S, 21:25:41.8 E). The layout pattern with antenna labels are show
in Fig. \ref{fig:AntPos}. The array consists of a 112-element grid and 16 "outriggers" used primarily to aid calibration.  In the bottom panel, the two baselines (5, 77) and (40, 85), marked in red, are an example of equivalent pair. We denote a equivalency-class of baselines in the PAPER grid by their separations, in this case  \{1,0\}, for the
antennas are separated by 1 unit east and 0 unit north. Similarly,
the baselines marked in blue and yellow are respectively examples
of \{1,1\} and \{1, -1\}.
Note \{1,0\} and \{-1, 0\} for example are the same class and should
not be counted twice. Antennas in purely north-south baselines
are close (4m), and hence these baselines are not suitable
for cross-multiplication due to cross-talks. On the other hand, the close North-South separation means that classes such as \{1,0\} and  \{1,1\} are expected to be near-equivalent. The PAPER-64 analysis of \cite{Ali2015} used three classes of baselines, the PAPER-128
equivalent of which are 
\{2, 0\}, \{2, 1\} and \{2, -1\}. There each of these classes
of baselines are cross multiplied within itself. This paper provides the method for inter-class multiplications. To do so we will use the short hand notation $\{m,n\}:\{m',n'\}$ to denote a pair of baseline classes to be cross-multiplied.

@@

\subsection{Rough Idea \label{sec:tracks}}

Given a point source on the sky, each baseline maps to a point in $uv$ plane. As the Earth rotates with respect to
the source, the points trace out tracks in the $uv$ plane. 
We show in Fig. \ref{fig:Tracks} $uv$ tracks of PAPER-128 over 4.8 hours, at 0.15GHz, for a source that passes through zenith.

@@

Roughly speaking, we can identify
redundancy of near-equivalent baselines as crossings
of the $uv$ tracks. As we see in Fig. \ref{fig:Tracks}, there are many such crossings. However, crossing tracks do not imply perfect redundancy in the case of drift-scan arrays. In fact, track-crossing is
not accurate enough for time offset determination, nor can it give estimate of the degree of redundancy.  The reason lies in the finite size of the beams and the nature of a drift-scan measurement. This shall become evident in the next section, after we develop a more general formalism that accounts for the point spread function of the finite beams, and estimate the degree of redundancy for general combination of baselines at a general time-offset. We show sample beams of HERA and PAPER antennas in Fig. \ref{fig:Beam} for reference.

@@

\caption{Example beam response of HERA (left) and PAPER (right) antennas. Both beams are Stokes $I$ polarization antenna voltage beams, at frequency of $\nu=150\text{MHz}$ and normalized to $1$ at zenith. Notice $A$ in this paper is a ``baseline's beam'', equivalent to squares of the antenna voltage beams shown here. The circles centered around zenith (center of beam) here are
spaced 10 degrees apart. \label{fig:Beam}}
\end{figure}

\subsection{Formalism}
In this section we formulate the relation between 
the product of visibilities from two arbitrary baselines and the power spectrum.

We begin with the visibility as commonly defined in the literature (e.g.
\citealt{first-paper}): 
\begin{equation}
\begin{aligned}V_{\nu}(\boldsymbol{b}) & =\int d\Omega A_{\nu}(\hat{\boldsymbol{s}})\phi(\nu)I_{\nu}(\hat{\boldsymbol{s}})\exp\left(-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right),\\
 & \approx\frac{2k_{B}}{\lambda^{2}}\int d\Omega A_{\nu}(\hat{\boldsymbol{s}})\phi(\nu)T(\hat{\boldsymbol{s}})\exp\left(-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right),
\end{aligned}
\label{eq:Vis1}
\end{equation}
Here $k_B$ is Boltzmann's constant, $\lambda$ is the mean wavelength, $\boldsymbol{b}$ is the baseline
length, $\hat{\boldsymbol{s}}$ and $\Omega$ are a direction in the
sky and its corresponding solid angle. Inside the integral we have $\phi(\nu)$ the frequency bandpass profile, $A_{\nu}$ the (frequency
dependent) primary beam, and $I$ the specific intensity, which has been
related to $T$, the brightness temperature in the Rayleigh-Jeans
limit. The beam reception pattern $A$ is dimensionless,
normalized to 1 at its peak (zenith), and we assume it to be the same
for all baselines. In practice,
power spectrum measurements are typically taken from a few ten MHz centered around the corresponding redshift of interest (e.g. 150 MHz for z=9.5). Note that in Eq. \ref{eq:Vis1}, no flat-sky approximation has yet been made; the angular integral is performed over the dome.

Delayed-transformed visibility has been widely adopted in recent years since it has been shown that foregrounds are isolated in delay space. We define the delay-transformed visibility by Fourier transforming visibility along the frequency axis\citep{delay-transform} as:
\small
\begin{equation}
\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left(-2\pi i\nu\tau\right),\\
 & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{\boldsymbol{s}},\nu)\phi(\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}+\tau\right)\right]. 
\end{aligned}
\label{eq:Vb1}
\end{equation}
\normalsize
Here the delay $\tau$ is the Fourier variable. Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
an integral over observational coordinates $\hat{\boldsymbol{s}}$ and $\nu$. Ultimately,
we would like to relate the data, collected with coordinates $\hat{\boldsymbol{s}}$
and $\nu$, to the power spectrum, written with 
$\boldsymbol{r}$ and $\boldsymbol{k}$, the cosmological position coordinate and wavenumber. To do so, we start by noticing that
\[
\begin{aligned}r & =\frac{c}{H_{0}}\int_{0}^{z}\frac{dz'}{E(z')},\\
 & \approx\frac{c}{H_{0}}\int_{0}^{z_{0}}\frac{dz'}{E(z')}-\frac{c(1+z)^{2}}{\nu_{21}H_{0}E(z)}\left(\nu-\nu_{0}\right),\\
 & \equiv X-Y\Delta\nu,
\end{aligned} \label{eq:r}
\]
where $\nu_{21}=1420$MHz is the 21\,cm transition rest frequency, $\nu_{0}$
a reference central frequency with corresponding redshift $z_{0}$,
and 
\[
E(z)=\sqrt{\Omega_{m}(1+z)^{3}+\Omega_{\Lambda}}.
\]
Inverting for $\nu$:
\begin{equation}
\nu=\frac{X-r}{Y}+\nu_{21}.\label{eq:nur}
\end{equation}

Further we shall invoke the flat-sky limit, with the $z$ direction points to zenith, and $x$ and $y$ point east and north, respectively:
\begin{equation}
(r_{x},r_{y})=(X\hat{\boldsymbol{s}}_{x},X\hat{\boldsymbol{s}}_{y}). 
\end{equation}
In the same limit $(Xk_{x},Xk_{y},Yk_{z})=\frac{2\pi}{c}(b_{x},b_{y},\tau)$ relates
the cosmological $k$ to the observational coordinates. For a more rigorous treatment without the flat sky limit, see Appendix [ADRIAN].

We can rewrite the delayed-transformed visibility as 
\small
\[
\begin{aligned}V(\boldsymbol{b},\tau) & =\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(\boldsymbol{r})\phi(r)T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\hat{\boldsymbol{r}}+\tau\right)\nu_{r}\right]\end{aligned},
\]
\normalsize
 where $d\nu=-dr/Y$ and $d^{3}r=-X^{2}Yd\Omega d\nu$. 
We have written $\nu_{r}$ as a reminder that $\nu$ and $r$ are related
by Eq. \eqref{eq:nur}.

Most existing visibility based power-spectrum pipelines for redundant drift-scan  arrays relate the power-spectrum to the conjugate square of the visibilities \citep{first-paper, paper32, Ali2015}. We would like to generalize such relations by relating the power-spectrum to the product of two visibilities from two arbitrary baselines and time-offsets. 
The beam pattern of a baseline shifts relative to the sky as the Earth rotates. Here we choose to fix the sky, and denote the rotated coordinates
of the beam pattern with the 3 dimensional rotation operator $\Gamma$:
\small
\[
\begin{aligned}V(\boldsymbol{b'},\tau') & =\frac{2k_{B}}{\lambda^{2}}\int\frac{d^{3}r}{X^{2}Y}A(\Gamma\boldsymbol{r})\phi(r)T(\boldsymbol{r})\exp\left[-2\pi i\left(\frac{\boldsymbol{b}}{c}\cdot\Gamma\hat{\boldsymbol{r}}+\tau'\right)\nu_{r}\right]\end{aligned}.
\]
\normalsize
With implicit bounds of integrals from $-\infty$ to $\infty$, we have:
\begin{widetext}
\begin{equation}
\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau')\rangle\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\langle T^{*}(\boldsymbol{r})T(\boldsymbol{r'})\rangle\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}\frac{d^{3}\kappa'}{(2\pi)^{3}}\langle T^{*}(\boldsymbol{\kappa})T(\boldsymbol{\kappa'})\rangle e^{-i(\boldsymbol{\kappa}\cdot \boldsymbol{r}-\boldsymbol{\kappa'}\cdot\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}\int d^{3}rd^{3}r'\left(\int\frac{d^{3}\kappa}{(2\pi)^{3}}P(\kappa)e^{-i\boldsymbol{\kappa}\cdot(\boldsymbol{r}-\boldsymbol{r'})}\right)A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & \approx\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}rd^{3}r'\delta_{D}^{(3)}(\boldsymbol{r}-\boldsymbol{r'})A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r'})\Phi_{b,\tau}(\boldsymbol{r},\Gamma \boldsymbol{r'}),\\
 & =\left(\frac{2k_{B}}{X^{2}Y\lambda^{2}}\right)^{2}P(k_{b,\tau})\int d^{3}r|A^{*}(\boldsymbol{r})A(\Gamma \boldsymbol{r})||\phi(\nu_{r})|^{2}\exp\left[-i2\pi\frac{\nu_{r}}{c}\left(\hat{\boldsymbol{r}}\cdot\boldsymbol{b}-\Gamma \hat{\boldsymbol{r}}\cdot\boldsymbol{b'}\right)\right],\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma \hat{\boldsymbol{s}},\nu)||\phi(\nu)|^{2}\exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)\right],
\end{aligned}
\label{eq:main}
\end{equation}
